---
title: "684 Midterm Project Proposal"
subtitle: "Allstate Purchase Prediction"
author: "Mengyun Li"
date: "November 13, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("dplyr")
```

# 1. Introduction

As a customer shops an insurance policy, he/she will receive a number of quotes with different coverage options before purchasing a plan. This is represented in this challenge as a series of rows that include a customer ID, information about the customer, information about the quoted policy, and the cost. The task is to predict the purchased coverage options using a limited subset of the total interaction history.

# 2. Data and Method

## 2.1 Data Source

```{r}
#Data clean
data_test <- read.csv("test_v2.csv")
data_test <- na.omit(data_test)
data_new <- data_test[5:12004, ] 
head(data_new, n=5)
```

## 2.2 Variable Discriptions
customer_ID - A unique identifier for the customer  
shopping_pt - Unique identifier for the shopping point of a given customer  
record_type - 0=shopping point, 1=purchase point  
day - Day of the week (0-6, 0=Monday)  
time - Time of day (HH:MM)  
state - State where shopping point occurred  
location - Location ID where shopping point occurred  
group_size - How many people will be covered under the policy (1, 2, 3 or 4)  
homeowner - Whether the customer owns a home or not (0=no, 1=yes)  
car_age - Age of the customer¡¯s car  
car_value - How valuable was the customer¡¯s car when new  
risk_factor - An ordinal assessment of how risky the customer is (1, 2, 3, 4)  
age_oldest - Age of the oldest person in customer's group  
age_youngest - Age of the youngest person in customer¡¯s group  
married_couple - Does the customer group contain a married couple (0=no, 1=yes)  
C_previous - What the customer formerly had or currently has for product option C (0=nothing, 1, 2, 3,4)  
duration_previous -  how long (in years) the customer was covered by their previous issuer
A,B,C,D,E,F,G - the coverage options  
cost - cost of the quoted coverage options  

# 3. Exploratory Data Analysis


## 3.1 Visuallization of the Data


